```{r}
library(readr)
library(dplyr)
library(lfe)
library(stargazer)
library(jsonlite)
library(tidyr)
library(stringr)
library(purrr)
library(ggplot2)
library(lubridate)
library(data.table)

# install.packages("tidyr","ggplot2","lubridate","stargazer")
```

```{r}
stage_df <- read_csv("results_df_2021_2023.csv") %>%
  mutate(time = hms(time)) %>%
  bind_rows(read_csv("results_df_2018_2020.csv")%>%mutate(time = hms(time))) %>%
  filter(
    !is.na(team_name),
    uci_tour %in% c("UCI Worldtour","UCI ProSeries"),
    #is_one_day_race, # For now - stage races more complex
    #stage_type == "RR",
    # status == "DF"
  ) %>%
  mutate(
    total_uci_points = uci_points + coalesce(gc_uci_points,0) + coalesce(points_uci_points,0) + coalesce(kom_uci_points,0) + coalesce(youth_uci_points,0),
    race_year = year(date), # assuming no races over new years
    vm_per_km = vertical_meters/distance, # for potential profile grouping
    abandon = status != "DF",
    stage_number = as.numeric(str_extract(stage_url, "(?<=stage-)\\d+"))
  ) %>%
  group_by(race_name,race_year) %>%
  mutate(race_date = min(date))


# Create a list of unique race dates
unique_dates <- unique(stage_df$race_date)

# code for the same for riders is in old.rmd

stage_df <- unique_dates %>%
  map(~{
    stage_df %>%
      filter(race_date == .x) %>%  # Focus on the current race date
      left_join(
        stage_df %>%
          filter(
            race_date < .x & race_date >= (.x %m-% months(12)) # Within the last 12 months
          ) %>%
          group_by(team_name) %>%
          summarize(team_points_l12m = sum(total_uci_points, na.rm = TRUE), .groups = "drop"),
        by = "team_name"
      ) %>%
      mutate(period = .x)  # Add a column for the current race date
  }) %>%
  bind_rows()
```

I want to examine stage race results, comparing teams that lost riders vs. those that didn't. ODR crashes are too likely to impact teammates outside of the rider abandoning.
- team-race level for only stage race
vars:
- abandons
- % of race missed by abandons - 1 would mean the equivalent of 1 rider missing for the whole race. So this can go higher than 1
- extra reg vars -- team_ppr etc

```{r}
# Create a list of unique race dates
unique_dates <- unique(stage_df$race_date)

# code for the same for riders is in old.rmd

stage_race_reg_df <- unique_dates %>%
  map(~{
    stage_df %>%
      filter(race_date == .x) %>%  # Focus on the current race date
      left_join(
        stage_df %>%
          filter(
            race_date < .x & race_date >= (.x %m-% months(12)) # Within the last 12 months
          ) %>%
          group_by(team_name) %>%
          summarize(team_points_l12m = sum(total_uci_points, na.rm = TRUE), .groups = "drop"),
        by = "team_name"
      ) %>%
      mutate(period = .x)  # Add a column for the current race date
  }) %>%
  bind_rows() %>%
  group_by(team_name, race_name, race_year, team_points_l12m) %>%
  summarize(
    uci_points = sum(total_uci_points),
    is_one_day_race = max(is_one_day_race),
    pct_missed_by_abandon_stage_race = sum((max(stage_number) - ifelse(abandon,stage_number,max(stage_number)))/max(stage_number)),
    max_stage = max(stage_number),
    riders_finished = sum(coalesce(stage_number,-1) == coalesce(max_stage,-1) & status == "DF"),
    race_date = min(date)
  ) %>%
  group_by(race_name,race_year) %>%
  mutate(
    team_ppr = uci_points/riders_finished,
    team_points = sum(uci_points),
    total_ppr = sum(uci_points)/sum(riders_finished),
    team_ppr_adj = team_ppr/total_ppr
  ) %>%
  ungroup() %>%
  group_by(team_name,race_year) %>%
  mutate(
    team_yearly_points = sum(uci_points)
  ) %>%
  filter(!is.na(team_ppr_adj),
         is.finite(team_ppr_adj))
```


How did the team's points per finishing rider differ when riders abandoned in a stage race?
```{r}
m1 = stage_race_reg_df %>% felm(team_ppr_adj ~ pct_missed_by_abandon_stage_race,data=.)
m2 = stage_race_reg_df %>% felm(team_ppr_adj ~ pct_missed_by_abandon_stage_race | race_name + race_year,data=.)
m3 = stage_race_reg_df %>% felm(team_ppr_adj ~ pct_missed_by_abandon_stage_race + team_points_l12m,data=.)
m4 = stage_race_reg_df %>% felm(team_ppr_adj ~ pct_missed_by_abandon_stage_race + team_points_l12m | race_name + race_year,data=.)

stargazer(m1,m2,m3,m4, type='text')
```

Stable across years?
```{r}
felm(team_ppr_adj ~ pct_missed_by_abandon_stage_race * factor(race_year), data = stage_race_reg_df) %>% stargazer(type='text')
```

So we see that teams do worse, per finishing rider, when they lose riders. This effect diminishes a bit when we control for a team's prior success - meaning (that better teams are more likely to lose riders?) but it is always between -0.05 and -0.11, so a 5 to 11% drop in performance per rider missing. This is not quite a marginal number, as some teams have multiple riders missing, and we'd expect that losing the second rider harms you more than losing the first rider. Something like an average treatment effect across actual losses of rider (ie teams don't often finish with just 1 rider)

Alternatively: we look at teams that had abandons near the middle of the race and look at how they did before/after. To do this:
* Build a virtual points stat: if the race ended now, what would that riders' points be? Adjust stage points to reflect how much of the race has happened
* Filter for races where riders abandoned all on one day (for clear effects)
* get all non-abandoning riders and sum their virtual points for each stage since crash along with actual points at the end
* Event study

Also can just filter to times where this happened near the middle and check points before/points after
* Also maybe points from the day after, in case the DNF itself caused issues

Can also just look at GC placings + stage wins/points. Will start with that

```{r}
stage_race_pre_post_df = stage_df %>%
  filter(!is_one_day_race) %>%
  group_by(race_name,race_year) %>%
  mutate(last_stage = max(stage_number),
         stage_1t = ceiling(last_stage/3)) %>%
  group_by(race_name,race_year,rider_name) %>%
  arrange(date) %>%
  mutate(
    win = coalesce(ifelse(rank == 1, 1, 0),0),
    cumulative_stage_points = cumsum(uci_points),
    cumulative_wins = cumsum(ifelse(rank == 1, 1, 0)),
    abandon_stage = max(ifelse(status != c("DF"),stage_number,0),na.rm=T),
    abandon_stage = ifelse(abandon_stage == 0, NA_integer_,abandon_stage)
  ) %>%
  ungroup() %>%
  group_by(stage_url) %>%
  mutate(win_adj = win/sum(win,na.rm=T)) %>%
  ungroup() %>%
  group_by(race_name,race_year,team_name) %>%
  mutate(first_abandon_stage = min(abandon_stage,na.rm = T),
         last_abandon_stage = max(abandon_stage,na.rm = T),
         abandoned = !is.na(abandon_stage),
         team_abandoned = !is.infinite(first_abandon_stage)) %>%
  mutate(
    grouping = case_when(
      !team_abandoned ~ "No abandons",
      (first_abandon_stage > stage_1t & last_abandon_stage <= last_stage - stage_1t) ~ "Abandons only in middle third",
      T ~ "Abandons outside middle third"
    )
  ) %>%
  ungroup() %>%
  group_by(race_name,race_year,team_name, grouping) %>%
  summarize(
    best_placed_gc_1t = min(ifelse(stage_number == stage_1t,gc_rank,NA_integer_),na.rm=T),
    best_placed_gc_1t_finishers = min(ifelse(stage_number == stage_1t & !abandoned,gc_rank,NA_integer_),na.rm=T),
    best_placed_gc_finish = min(ifelse(stage_number == last_stage,gc_rank,NA_integer_),na.rm=T),
    avg_placed_gc_1t = mean(ifelse(stage_number == stage_1t,gc_rank,NA_integer_),na.rm=T),
    avg_placed_gc_1t_finishers = mean(ifelse(stage_number == stage_1t & !abandoned,gc_rank,NA_integer_),na.rm=T),
    avg_placed_gc_finish = mean(ifelse(stage_number == last_stage,gc_rank,NA_integer_),na.rm=T),
    stage_wins_per_stage_1t = sum(ifelse(stage_number <= stage_1t,win_adj,NA_integer_),na.rm=T)/max(stage_1t),
    stage_wins_per_stage_1t_finishers = sum(ifelse(stage_number <= stage_1t & !abandoned,win_adj,NA_integer_),na.rm=T)/max(stage_1t),
    stage_wins_per_stage_3t = sum(ifelse(stage_number > last_stage - stage_1t,win_adj,NA_integer_),na.rm=T)/max(stage_1t),
    stages_1t = n_distinct(ifelse(stage_number <= stage_1t,stage_number,NA_integer_),na.rm=T),
    stages_3t = n_distinct(ifelse(stage_number > last_stage - stage_1t,stage_number,NA_integer_),na.rm=T),
    stage_wins_1t = sum(ifelse(stage_number <= stage_1t,win_adj,NA_integer_),na.rm=T),
    stage_wins_3t = sum(ifelse(stage_number > last_stage - stage_1t,win_adj,NA_integer_),na.rm=T),
  )
```

```{r}
stage_race_pre_post_df %>%
  filter(!is.infinite(best_placed_gc_1t),
         !is.infinite(best_placed_gc_1t_finishers)) %>%
  ungroup() %>%
  select(-c(race_name,team_name)) %>%
  group_by(grouping) %>%
  summarise(across(everything(), mean))
```

```{r}
stage_race_pre_post_df %>%
  filter(!is.infinite(best_placed_gc_1t),
         !is.infinite(best_placed_gc_1t_finishers)) %>%
  ungroup() %>%
  select(-c(race_name,team_name)) %>%
  group_by(grouping) %>%
  summarise(across(everything(), sum))
```

# Break teammates
* Filter for breakaway wins
* Did people with teammates do better?

We'd expect this would show positive even with no teammate effect. A team's second best rider is a signal of the quality of their first best rider (mathematically, but also teams in cycling have very different budgets and correspondingly different average rider qualitites). So a team with 2 riders in a break is more likely to have a stronger 1st rider than a team with 1 rider. There's also interfering effects here. For example, in a stage race, worse teams may be more keen to get riders in breakaways while better teams save riders' energy or keep them back to (hypothetically) help keep their leaders out of trouble.

However it's still worth seeing if we can find any useful signal.

```{r}
gap_length_seconds = 30
max_break_size = 15

first_gaps = stage_df %>%
  mutate(time = hms(time)) %>%
  # Find first gap bigger than X seconds
  group_by(stage_url) %>%
  mutate(time_gap = lead(time,order_by=rank) - time) %>%
  filter(time_gap > gap_length_seconds) %>%
  summarize(first_rank_with_gap = min(rank))

stage_df %>%
  merge(first_gaps, by="stage_url",all.x=T) %>%
  # mutate(stage_profile = NA) %>% # fill out - we want to break out by flat, hilly, mountains. and maybe by finish profile if poss
  # Only the break in valid stages
  filter(first_rank_with_gap <= max_break_size,
         rank <= first_rank_with_gap) %>%
  group_by(stage_url,team_name) %>%
  summarize(teammates_in_first_break = n(),
            best_placed_in_team = min(rank),
            avg_placed_in_team = mean(rank)) %>%
  group_by(stage_url) %>%
  mutate(break_size = sum(teammates_in_first_break)) %>%
  # Controlling for break size as larger break means more likely to have teammates and also lower average best placed. Maybe this is not the best way to do it though? Too tired to think
  # Best placed in team has a nonlinear relationship with break size?
  lm(avg_placed_in_team ~ teammates_in_first_break + break_size,data=.)

%>%
  summary()
```

Can placebo here. What is the lift if given a random teammate?

General hypotheses:

* Trains to keep riders at the front do basically nothing, either for GC or sprints
* Mountain domestiques to very little, especially for underdogs
* Attacking in turns is more effective relative to leadouts
* One good leadout man is better than a team of middling leadout men
* Breakaways are relatively underused. EG tim declerq should be going for breaks more often

Talk about pidcock/rodriguez in tour doc. These are oversensationalized but they often contain some truth, like pidcock actually beefing with management. What benefit could he have possibly given carlos??

How can riders use game theory more effectively?

https://www.youtube.com/watch?v=SQ01eBEJ1Tk - 104:30 on unlocking riders

* Fakes:
  * Riders should want opposing riders to not know when they will attack. One way: launch fake attacks, or ride so that lead up to an attack looks just like normal riding
  * Or: some portion of attacks should fizzle early. That way teams will not chase so hard
* 
