```{r}
library(readr)
library(dplyr)
library(lfe)
library(stargazer)
library(jsonlite)
library(tidyr)
library(stringr)
library(purrr)
library(ggplot2)
library(lubridate)
```

```{r}
stage_df <- read_csv("results_df_2021_2023.csv") %>%
  filter(
    !is.na(team_name),
    uci_tour %in% c("UCI Worldtour","UCI ProSeries"),
    #is_one_day_race, # For now - stage races more complex
    #stage_type == "RR",
    # status == "DF"
  ) %>%
  mutate(
    total_uci_points = uci_points + coalesce(gc_uci_points,0) + coalesce(points_uci_points,0) + coalesce(kom_uci_points,0) + coalesce(youth_uci_points,0)
  )

url_lookup = read_csv("url_lookup_2021_2023.csv")
  
crash_df = read_csv("crash_df_2021_2023.csv") %>%
  mutate(riders = map(riders, ~ {
    # Step 2: Remove brackets but preserve content inside
    list_string <- str_remove_all(.x, "\\[|\\]")
    
    # Step 3: Split by ', ' while preserving apostrophes inside names
    str_split(list_string, "', '", simplify = TRUE) %>%
      str_remove_all("'")  # Remove remaining single quotes
  })) %>%
  unnest(riders) %>%
  rename(stage_url_2 = stage_url) %>%
  merge(url_lookup,by="stage_url_2") %>%
  select(-stage_url_2) %>%
  rename(rider_name = riders)
```

Are crashes more likely reported for high profile riders? Although biased as crashing will reduce total points...
```{r}
rider_year_points = stage_df %>%
  group_by(rider_name) %>%
  summarize(total_points = sum(total_uci_points),
            total_stages = n())

crash_df %>%
  group_by(rider_name) %>%
  summarize(crashes = n()) %>%
  merge(rider_year_points, by = "rider_name", all.y = TRUE) %>%
  mutate(crashes = coalesce(crashes, 0),
         crashes_per_stage = crashes/total_stages) %>%
  ggplot(aes(x = total_points, y = crashes_per_stage)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE, color = "blue") # Adds a linear trendline
```

What about high profile races?
```{r}
race_points = stage_df %>%
  group_by(stage_url) %>%
  summarize(total_points = sum(total_uci_points))

crash_df %>%
  group_by(stage_url) %>%
  summarize(crashes = n()) %>%
  merge(race_points,by="stage_url",all.y=T) %>%
  mutate(crashes = coalesce(crashes,0)) %>%
  ggplot(aes(x=total_points,y=crashes)) + geom_point() +
  geom_smooth(method = "lm", se = TRUE, color = "blue") # Adds a linear trendline
```
Get the UCI points given out per rider for each race, filtered:
* One day road races (for now, as tours would require to also count following stages as DNFs)
* World or pro tour
* Finishes
```{r}
stage_df %>%
  filter(
    !is.na(team_name),
    uci_tour %in% c("UCI Worldtour","UCI ProSeries"),
    # is_one_day_race,
    # stage_type == "RR",
    status == "DF"
  ) %>%
  summarize(
    uci_points_2 = sum(uci_points) + sum(gc_uci_points,na.rm=T) + sum(points_uci_points,na.rm=T) + sum(kom_uci_points,na.rm=T) + sum(youth_uci_points,na.rm=T),
    uci_points_per_finish = sum(uci_points_2)/n(),
    gc_points = sum(gc_uci_points,na.rm=T)
  )
```

```{r}
stage_df %>% 
  mutate(
    total_uci_points = uci_points + coalesce(gc_uci_points,0) + coalesce(points_uci_points,0) + coalesce(kom_uci_points,0) + coalesce(youth_uci_points,0)
  ) %>%
  arrange(desc(uci_points_2)) %>% select(stage_url,uci_points_2)
```

Get crash DNFs
```{r}
crash_df_temp = crash_df %>%
  group_by(rider_name,stage_url) %>%
  summarize() %>%
  mutate(crash = T)
  
team_race_df = stage_df %>%
  merge(crash_df_temp,by=c("stage_url","rider_name"),all.x=T) %>%
  mutate(crash_dnf = case_when(crash & status == "DNF"~T,T~F)) %>%
  group_by(team_name, stage_url, race_name, date) %>%
  summarize(
    dnfs = sum(crash_dnf),
    uci_points = sum(uci_points)
  ) 

team_race_df %>%
  felm(uci_points ~ dnfs | race_name + team_name,data=.) %>%
  stargazer(type='text')
```

Get the loss in UCI points per DNF
```{r}
pcs_2023_df %>%
  filter(
    !is.na(team_name),
    uci_tour %in% c("UCI Worldtour","UCI ProSeries"),
    # is_one_day_race,
    stage_type == "RR"
  ) %>%
  group_by(team_name, stage_url, race_name, date) %>%
  summarize(
    dnfs = sum(status == "DNF"),
    uci_points = sum(uci_points)
  ) %>%
  felm(uci_points ~ dnfs | race_name + team_name,data=.) %>%
  stargazer(type='text')
```


Create clean df for the regressions:
- team-race level
-- including stage races summed up
vars:
- one day race
- elevation/distance
- crash dnfs
- % of race missed by crash dnfs
- total uci points
- # riders finished
- total race points

Plus: extra reg vars

```{r}
crash_df_temp = crash_df %>%
  group_by(rider_name,stage_url) %>%
  summarize(last_crash_kms = max(km_to_go)) %>%
  mutate(crash = T)

race_temp_df = stage_df %>%
  mutate(vm_per_km = vertical_meters/distance,
         race_year = year(date)) %>%
  group_by(stage_url, race_name, race_year, vertical_meters, distance) %>%
  summarize() %>%
  group_by(race_name,race_year) %>%
  summarize(
    vm_per_km = sum(vertical_meters)/sum(distance)
  )

rider_year_temp_df = stage_df %>%
  mutate(race_year = year(date)) %>%
  filter(!is.na(race_year)) %>%
  group_by(rider_name,race_year) %>%
  summarize(
    rider_ppr = sum(uci_points)/n()
  )

reg_df = stage_df %>%
  mutate(race_year = year(date)) %>%
  merge(crash_df_temp,by=c("stage_url","rider_name"),all.x=T) %>%
  merge(rider_year_temp_df,by=c("rider_name","race_year"),all.x=T) %>%
  mutate(crash_dnf = case_when(crash & status == "DNF"~T,T~F),
         crash_dnf_pct_missed = case_when(crash & status == "DNF"~coalesce(last_crash_kms*1.00/distance,0.5),T~0),
         stage_number = as.numeric(str_extract(stage_url, "(?<=stage-)\\d+"))) %>%
  merge(race_temp_df,by=c('race_name','race_year')) %>%
  group_by(team_name, race_name, race_year,vm_per_km) %>%
  summarize(
    dnfs = sum(crash_dnf),
    uci_points = sum(uci_points + coalesce(gc_uci_points,0) + coalesce(points_uci_points,0) + coalesce(kom_uci_points,0) + coalesce(youth_uci_points,0)),
    is_one_day_race = max(is_one_day_race),
    pct_missed_by_crash_dnfs = case_when(
      max(is_one_day_race) == 1 ~ sum(crash_dnf_pct_missed),
      T ~ sum((max(stage_number) - ifelse(crash_dnf,stage_number,0))/max(stage_number))
    ),
    lost_rider_ppr_weighted = case_when(
      max(is_one_day_race) == 1 ~ sum(crash_dnf_pct_missed*rider_ppr)/sum(crash_dnf_pct_missed),
      T ~ sum(((max(stage_number) - ifelse(crash_dnf,stage_number,0))/max(stage_number))*rider_ppr)/sum((max(stage_number) - ifelse(crash_dnf,stage_number,0))/max(stage_number))
    ),
    max_stage = max(stage_number),
    riders_finished = sum(coalesce(stage_number,-1) == coalesce(max_stage,-1) & status == "DF")
  ) %>%
  group_by(race_name,race_year) %>%
  mutate(
    ppr = uci_points/riders_finished,
    tp = sum(uci_points),
    tppr = sum(uci_points)/sum(riders_finished),
    ppr_over_tppr = ppr/tppr
  ) %>%
  ungroup() %>%
  group_by(team_name,race_year) %>%
  mutate(
    team_tp = sum(uci_points)
  )
```

How 
```{r}
reg_df %>%
  filter(!is.na(ppr_over_tppr),
         is.finite(ppr_over_tppr),
         !is.na(dnfs)) %>%
  felm(ppr_over_tppr ~ dnfs | tp ,data=.) %>%
  stargazer(type='text')
```

```{r}
reg_df %>%
  filter(!is.na(ppr_over_tppr),
         is.finite(ppr_over_tppr),
         !is.na(dnfs)) %>%
  felm(ppr_over_tppr ~ pct_missed_by_crash_dnfs + tp + team_tp + lost_rider_ppr_weighted,data=.) %>%
  stargazer(type='text')
```

```{r}
reg_df %>%
  filter(!is.na(ppr_over_tppr),
         is.finite(ppr_over_tppr),
         !is.na(dnfs)) %>%
  felm(uci_points ~ pct_missed_by_crash_dnfs*tp + team_tp + lost_rider_ppr_weighted,data=.) %>%
  stargazer(type='text')
```
