# Load packages
```{r}
library(readr)
library(dplyr)
library(lfe)
library(stargazer)
library(jsonlite)
library(tidyr)
library(stringr)
library(purrr)
library(ggplot2)
library(lubridate)
library(data.table)

# install.packages("tidyr","ggplot2","lubridate","stargazer")
```

# TODOs
I think we're upwardly biased, because one team doing well means other teams do less well. 1/20 or so, so maybe not important?
update with full dataset and go through findings
make sure i'm not including points of non-finishing riders in PPR calcs
change varname uci_points, confusing - sometimes stage sometimes combined
change gross code highlighted below
update breakaway tagging to be more complex
 - we're not including times where there was a solo break with a group behind
 - won_how is pretty good - could switch to that

# Set up base dataset stage_df
```{r}
stage_df <- read_csv("results_df_2021_2023.csv") %>%
  mutate(time = hms(time)) %>%
  # bind_rows(read_csv("results_df_2018_2020.csv")%>%mutate(time = hms(time))) %>%
  filter(
    !is.na(team_name),
    uci_tour %in% c("UCI Worldtour","UCI ProSeries"),
    #is_one_day_race, # For now - stage races more complex
    #stage_type == "RR",
    # status == "DF"
  ) %>%
  mutate(
    total_uci_points = uci_points + coalesce(gc_uci_points,0) + coalesce(points_uci_points,0) + coalesce(kom_uci_points,0) + coalesce(youth_uci_points,0),
    race_year = year(date), # assuming no races over new years
    vm_per_km = vertical_meters/distance, # for potential profile grouping
    abandon = status != "DF",
    stage_number = as.numeric(str_extract(stage_url, "(?<=stage-)\\d+"))
  ) %>%
  group_by(race_name,race_year) %>%
  mutate(race_date = min(date))

# Get a team's total points in the last 12 months before each race - to proxy for team strength

# Create a list of unique race dates
unique_dates <- unique(stage_df$race_date)

# code for the same for riders is in old.rmd

stage_df <- unique_dates %>%
  map(~{
    stage_df %>%
      filter(race_date == .x) %>%  # Focus on the current race date
      left_join(
        stage_df %>%
          filter(
            race_date < .x & race_date >= (.x %m-% months(12)) # Within the last 12 months
          ) %>%
          group_by(team_name) %>%
          summarize(team_points_l12m = sum(total_uci_points, na.rm = TRUE), .groups = "drop"),
        by = "team_name"
      ) %>%
      mutate(period = .x)  # Add a column for the current race date
  }) %>%
  bind_rows()
```

First question: how does losing a rider affect a team's performance?

Let's look at stage races for this. One-day races are tough because if a rider DNFs, they were likely caught in a crash, which could involve a teammate - meaning their results are hurt by the DNF more than just the loss of one teammate. For now, we'll look at all DNFs - later, we might try to isolate to just crash-related DNFs, which are more random and so should help us get a better understanding. I'll use pct_missed_by_abandon as my treatment variable - this is basically the stages missed by riders who abandoned the race divided by the number of stages in a race. So a value of 1 implies that the team missed the equivalent of 1 rider for every stage in the race - for example, by losing two riders for half the race each.

(The impact of teammates is almost definitely non-linear, but we'll ignore that for now)

I'll also use team_ppr_adj as my dependent variable. This is constructed by first summing all UCI points earned by the finishing riders of a team in the stage race (stage points as well as jersey points); then dividing those points by the number of finishing riders on the team; then dividing that number by the average points per finishing riders for the whole stage race. This is slightly convoluted, but it gives us a measure of a team's performance relative to the average performance, adjusted for how many riders finished the race - and it keeps scale across races with very different point totals, which is important as bigger races give a lot more UCI points out.

# Create regression df
```{r}
# code for the same for riders is in old.rmd

stage_race_reg_df <- stage_df %>%
  group_by(team_name, race_name, race_year, team_points_l12m) %>%
  summarize(
    uci_points = sum(total_uci_points),
    is_one_day_race = max(is_one_day_race),
    pct_missed_by_abandon_stage_race = sum((max(stage_number) - ifelse(abandon,stage_number,max(stage_number)))/max(stage_number)),
    max_stage = max(stage_number),
    riders_finished = sum(coalesce(stage_number,-1) == coalesce(max_stage,-1) & status == "DF"),
    race_date = min(date)
  ) %>%
  group_by(race_name,race_year) %>%
  mutate(
    team_ppr = uci_points/riders_finished,
    team_points = sum(uci_points),
    total_ppr = sum(uci_points)/sum(riders_finished),
    team_ppr_adj = team_ppr/total_ppr
  ) %>%
  ungroup() %>%
  group_by(team_name,race_year) %>%
  mutate(
    team_yearly_points = sum(uci_points)
  ) %>%
  filter(!is.na(team_ppr_adj),
         is.finite(team_ppr_adj))
```

# Run regressions
```{r}
m1 = stage_race_reg_df %>% felm(team_ppr_adj ~ pct_missed_by_abandon_stage_race,data=.)
m2 = stage_race_reg_df %>% felm(team_ppr_adj ~ pct_missed_by_abandon_stage_race | race_name + race_year,data=.)
m3 = stage_race_reg_df %>% felm(team_ppr_adj ~ pct_missed_by_abandon_stage_race + team_points_l12m,data=.)
m4 = stage_race_reg_df %>% felm(team_ppr_adj ~ pct_missed_by_abandon_stage_race + team_points_l12m | as.factor(race_name):as.factor(race_year),data=.)

stargazer(m1,m2,m3,m4, type='text')
```

# TODO below is unfinished and needs to be updated with full dataset

So we see that teams do worse, per finishing rider, when they lose riders. This effect diminishes a bit when we control for a team's prior success, meaning that historically worse teams are more likely to lose riders. And when I control for prior success AND WHAT but it is always between -0.05 and -0.11, so a 5 to 11% drop in performance per rider missing. This is not quite a marginal number, as some teams have multiple riders missing, and we'd expect that losing the second rider harms you more than losing the first rider. Something like an average treatment effect across actual losses of rider (ie teams don't often finish with just 1 rider)

I also find that these coefficients are stable across years - do I need this??
```{r}
felm(team_ppr_adj ~ pct_missed_by_abandon_stage_race * factor(race_year), data = stage_race_reg_df) %>% stargazer(type='text')
```

I'm worried though about selection bias here though. Riders might be more likely to abandon if their team is doing worse (and their cause is hopeless) or better (and the team doesn't need them). If this was happening in either direction, it would bias the results above. To investigate this, I've built an updated dataset with team results during stage races as well as at the end.

```{r}
# First - get UCI points given to current ranks, if the rider finishes with those ranks

# Step 1: Filter to final stage rows per race (assuming "stage" indicates stage number)
final_stage <- stage_df %>%
  group_by(race_name,race_year) %>%
  filter(stage_number == max(stage_number)) %>%
  ungroup()

# Step 2: Pivot final stage results to long format for classification ranks and points
# For ranks (e.g., gc_rank, kom_rank, etc.)
ranks <- final_stage %>%
  select(race_name,race_year, rider_name, gc_rank, kom_rank, points_rank, youth_rank) %>%
  pivot_longer(
    cols = c(gc_rank, kom_rank, points_rank, youth_rank),
    names_to = "classification",
    values_to = "jersey_rank"
  ) %>%
  mutate(classification = sub("_rank", "", classification))

# For points (e.g., gc_uci_points, kom_points, etc.)
points <- final_stage %>%
  select(race_name,race_year, rider_name, gc_uci_points, kom_uci_points, points_uci_points, youth_uci_points) %>%
  pivot_longer(
    cols = c(gc_uci_points, kom_uci_points, points_uci_points, youth_uci_points),
    names_to = "classification",
    values_to = "points"
  ) %>%
  mutate(classification = sub("(_uci)?_points", "", classification))

# Join the ranks and points together
final_long <- left_join(ranks, points, by = c("race_name","race_year", "rider_name", "classification"))

# Step 3: Summarize to get the maximum (final) points for each race, classification, and rank
points_table <- final_long %>%
  group_by(race_name,race_year, classification, jersey_rank) %>%
  summarise(final_points = max(points, na.rm = TRUE)) %>%
  ungroup() %>%
  filter(!is.na(jersey_rank))

# Step 4: Rejoin these final points onto the original stage-level dataframe
# Pivot stage_df to long format for classification ranks
stage_long <- stage_df %>%
  pivot_longer(
    cols = c(gc_rank, kom_rank, points_rank, youth_rank),
    names_to = "classification",
    values_to = "jersey_rank"
  ) %>%
  mutate(classification = sub("_rank", "", classification))

# Join with points_table based on race, classification, and rank
stage_long <- stage_long %>%
  left_join(points_table, by = c("race_name","race_year", "classification", "jersey_rank"))

# (Optional) Pivot back to wide format so each classification’s final points become separate columns.
# The new columns will be named final_points_gc, final_points_kom, etc.
stage_df_with_final_points <- stage_long %>%
  pivot_wider(
    names_from = classification,
    values_from = c(jersey_rank, final_points),
    names_sep = "_"
  )

# Now build the dataset
midpoint_df = stage_df_with_final_points %>%
  filter(!is_one_day_race) %>%
  group_by(race_name,race_year) %>%
  mutate(last_stage = max(stage_number),
         stage_1t = ceiling(last_stage/3)) %>%
  group_by(race_name,race_year,rider_name) %>%
  arrange(date) %>%
  mutate(
    win = coalesce(ifelse(rank == 1, 1, 0),0),
    cumulative_stage_points = cumsum(uci_points),
    cumulative_wins = cumsum(ifelse(rank == 1, 1, 0)),
    abandon_stage = max(ifelse(status != c("DF"),stage_number,0),na.rm=T),
    abandon_stage = ifelse(abandon_stage == 0, NA_integer_,abandon_stage),
    total_jersey_final_points = coalesce(final_points_gc,0) + coalesce(final_points_kom,0) + coalesce(final_points_points,0) + coalesce(final_points_youth,0)
  ) %>%
  ungroup() %>%
  group_by(stage_url) %>%
  mutate(win_adj = win/sum(win,na.rm=T)) %>%
  ungroup() %>%
  group_by(race_name,race_year,team_name) %>%
  mutate(first_abandon_stage = min(abandon_stage,na.rm = T),
         last_abandon_stage = max(abandon_stage,na.rm = T),
         abandoned = !is.na(abandon_stage),
         team_abandoned = !is.infinite(first_abandon_stage)) %>%
  mutate(
    grouping = case_when(
      !team_abandoned ~ "No abandons",
      (first_abandon_stage > stage_1t & last_abandon_stage <= last_stage - stage_1t) ~ "Abandons only in middle third",
      T ~ "Abandons outside middle third"
    )
  ) %>%
  ungroup() 

stage_race_pre_post_df = midpoint_df %>%
  group_by(race_name,race_year,team_name, team_points_l12m, grouping) %>%
  summarize(
    virtual_ppr_1t = mean(ifelse(stage_number == stage_1t,cumulative_stage_points + total_jersey_final_points/3,NA_integer_),na.rm=T),
    virtual_ppr_1t_finishers = mean(ifelse(stage_number == stage_1t & !abandoned,cumulative_stage_points + total_jersey_final_points/3,NA_integer_),na.rm=T),
    virtual_ppr_3t = mean(ifelse(stage_number == last_stage,cumulative_stage_points + total_jersey_final_points/3,NA_integer_),na.rm=T) - mean(ifelse(stage_number == last_stage - stage_1t & !abandoned,cumulative_stage_points,NA_integer_),na.rm=T),
    riders_1t = sum(ifelse(stage_number == stage_1t,1,0),na.rm=T),
    riders_3t = sum(ifelse(stage_number == last_stage,1,0),na.rm=T)
  ) %>%
  filter(!is.infinite(virtual_ppr_1t),
         !is.nan(virtual_ppr_1t),
         !is.nan(virtual_ppr_3t),
         !is.nan(riders_1t)) %>%
  ungroup() %>%
  group_by(race_name,race_year) %>%
  mutate(
    virtual_ppr_adj_1t = virtual_ppr_1t/(sum(virtual_ppr_1t*riders_1t)/sum(riders_1t)),
    virtual_ppr_adj_1t_finishers = virtual_ppr_1t_finishers/(sum(virtual_ppr_1t*riders_1t)/sum(riders_1t)),
    virtual_ppr_adj_3t = virtual_ppr_3t/(sum(virtual_ppr_3t*riders_3t)/sum(riders_3t))
  ) %>% 
  ungroup()
```

I measure results at the 1/3 mark of the race - the reason why will come clear below - and of course at the end. I get the part-way jersey results by assigning UCI points to each rider based on what they would earn if they ended the race with the same jersey placing. So first question now: do teams see more abandons after the first third of the race if they are doing poorly at that point?

```{r}
temp = stage_race_pre_post_df %>%
  filter(!is.infinite(virtual_ppr_adj_1t),
         !is.nan(virtual_ppr_adj_1t)) %>%
  mutate(abandons_23t = riders_1t - riders_3t) %>%
  ungroup() 


m1 = temp %>% felm(abandons_23t ~ virtual_ppr_adj_1t,data=.)
m2 = temp %>% felm(abandons_23t ~ virtual_ppr_adj_1t + team_points_l12m | as.factor(race_name):as.factor(race_year),data=.)

stargazer(m1,m2, type='text')
```

Clearly: yes. Top-performing teams can see adjusted PPRs of 5 or higher, while low-performing teams are usually close to 0; this coefficient would imply then that those top teams see, on average, roughly 1/3 fewer abandons than those bottom teams - even controlling for the race and the team's prior success.

Let's switch up these regressions a little - if we reverse the IV and DV, we get the impact that abandons in the last two thirds of the race have on results in the first third. Obviously there is no causal channel here (time flows forward) so this can help us understand the size of the error this selection bias could introduce into our earlier regression. Note - to compare apples to apples, we should divide the coefficients below by 3, as each abandon in the second 2/3 of the race will average about 1/3 of a lost rider-race.

```{r}
m1 = temp %>% felm(virtual_ppr_adj_1t ~ abandons_23t,data=.)
m2 = temp %>% felm(virtual_ppr_adj_1t ~ abandons_23t + team_points_l12m | as.factor(race_name):as.factor(race_year),data=.)

stargazer(m1,m2, type='text')
```

With just this selection effect, we'd expect an effect of about -0.5 in the earlier regression; or in other words, this bias swallows up a lot of the effect that we do observe. For our fullest regression, it eats away basically the whole thing! Takeaway: the impact of at least the first lost rider is likely quite small.

This is pretty surprising! We should interrogate it further. I'll use the same dataset for this: by comparing team results before and after abandons, we may be able to learn more.

For a clean-ish treatment, I focus on teams with abandons only in the middle third of the race. This gives us two thirds of data to compare for each group - one before any abandons and one after any abandons. The average team's PPR will be higher in the last third of a race because there are fewer riders; so I focus on how abandoned teams fare in the last third vs. the first third, relative to the same difference for teams with no abandons (I throw out teams with abandons outside the middle third of the race).

```{r}
temp = stage_race_pre_post_df %>%
  filter(!is.infinite(virtual_ppr_1t),
         !is.nan(virtual_ppr_1t),
         !is.nan(virtual_ppr_3t),
         grouping != 'Abandons outside middle third') %>%
  mutate(virtual_ppr_adj_increase = virtual_ppr_adj_3t - virtual_ppr_adj_1t,
         virtual_ppr_adj_increase_finishers = virtual_ppr_adj_3t - virtual_ppr_adj_1t_finishers,
         abandons = riders_1t - riders_3t)

m1 = temp %>% felm(virtual_ppr_adj_increase ~ abandons,data=.)
m2 = temp %>% felm(virtual_ppr_adj_increase ~ abandons + team_points_l12m | as.factor(race_name):as.factor(race_year),data=.)
m3 = temp %>% felm(virtual_ppr_adj_increase_finishers ~ abandons,data=.)
m4 = temp %>% felm(virtual_ppr_adj_increase_finishers ~ abandons + team_points_l12m | as.factor(race_name):as.factor(race_year),data=.)

stargazer(m1,m2,m3,m4, type='text')
```

This is all a bit of a mess. I don't think this is true. Guesses:
* Doing something wrong in the data setup
* Reg spec?

To test this, bring back in stage wins + GC placing

Can this be explained by changes in strategy to being more point focused?

Or end of stage races always different to start?

```{r}
temp = midpoint_df %>%
  group_by(race_name,race_year,team_name, team_points_l12m, grouping) %>%
  summarize(
    best_placed_gc_1t = min(ifelse(stage_number == stage_1t,jersey_rank_gc,NA_integer_),na.rm=T),
    best_placed_gc_1t_finishers = min(ifelse(stage_number == stage_1t & !abandoned,jersey_rank_gc,NA_integer_),na.rm=T),
    best_placed_gc_finish = min(ifelse(stage_number == last_stage,jersey_rank_gc,NA_integer_),na.rm=T),
    avg_placed_gc_1t = mean(ifelse(stage_number == stage_1t,jersey_rank_gc,NA_integer_),na.rm=T),
    avg_placed_gc_1t_finishers = mean(ifelse(stage_number == stage_1t & !abandoned,jersey_rank_gc,NA_integer_),na.rm=T),
    avg_placed_gc_finish = mean(ifelse(stage_number == last_stage,jersey_rank_gc,NA_integer_),na.rm=T),
    stage_wins_per_stage_1t = sum(ifelse(stage_number <= stage_1t,win_adj,NA_integer_),na.rm=T)/max(stage_1t),
    stage_wins_per_stage_1t_finishers = sum(ifelse(stage_number <= stage_1t & !abandoned,win_adj,NA_integer_),na.rm=T)/max(stage_1t),
    stage_wins_per_stage_3t = sum(ifelse(stage_number > last_stage - stage_1t,win_adj,NA_integer_),na.rm=T)/max(stage_1t),
    stages_1t = n_distinct(ifelse(stage_number <= stage_1t,stage_number,NA_integer_),na.rm=T),
    stages_3t = n_distinct(ifelse(stage_number > last_stage - stage_1t,stage_number,NA_integer_),na.rm=T),
    stage_wins_1t = sum(ifelse(stage_number <= stage_1t,win_adj,NA_integer_),na.rm=T),
    stage_wins_3t = sum(ifelse(stage_number > last_stage - stage_1t,win_adj,NA_integer_),na.rm=T),
    riders_1t = sum(ifelse(stage_number == stage_1t,1,0),na.rm=T),
    riders_3t = sum(ifelse(stage_number == last_stage,1,0),na.rm=T)
  ) %>%
  filter(grouping != 'Abandons outside middle third',
         !is.nan(riders_1t)) %>%
  mutate(abandons = riders_1t - riders_3t,
         best_placed_gc_increase = best_placed_gc_finish - best_placed_gc_1t,
         best_placed_gc_increase_finishers = best_placed_gc_finish - best_placed_gc_1t_finishers,
         stage_wins_per_stage_increase = stage_wins_per_stage_3t - stage_wins_per_stage_1t,
         stage_wins_per_stage_increase_finishers = stage_wins_per_stage_3t - stage_wins_per_stage_1t_finishers,
         avg_placed_gc_increase = avg_placed_gc_finish - avg_placed_gc_1t,
         avg_placed_gc_increase_finishers = avg_placed_gc_finish - avg_placed_gc_1t_finishers,
         ) %>%
  filter(is.finite(best_placed_gc_increase))

m1 = temp %>% felm(best_placed_gc_increase ~ abandons,data=.)
m2 = temp %>% felm(best_placed_gc_increase ~ abandons + team_points_l12m | as.factor(race_name):as.factor(race_year),data=.)
m3 = temp %>% felm(best_placed_gc_increase_finishers ~ abandons,data=.)
m4 = temp %>% felm(best_placed_gc_increase_finishers ~ abandons + team_points_l12m | as.factor(race_name):as.factor(race_year),data=.)
m5 = temp %>% felm(stage_wins_per_stage_increase ~ abandons,data=.)
m6 = temp %>% felm(stage_wins_per_stage_increase ~ abandons + team_points_l12m | as.factor(race_name):as.factor(race_year),data=.)
m7 = temp %>% felm(stage_wins_per_stage_increase_finishers ~ abandons,data=.)
m8 = temp %>% felm(stage_wins_per_stage_increase_finishers ~ abandons + team_points_l12m | as.factor(race_name):as.factor(race_year),data=.)

stargazer(m1,m2,m3,m4,m5,m6,m7,m8, type='text')
```

average placing?

```{r}
m1 = temp %>% felm(avg_placed_gc_increase ~ abandons,data=.)
m2 = temp %>% felm(avg_placed_gc_increase ~ abandons + team_points_l12m | as.factor(race_name):as.factor(race_year),data=.)
m3 = temp %>% felm(avg_placed_gc_increase_finishers ~ abandons,data=.)
m4 = temp %>% felm(avg_placed_gc_increase_finishers ~ abandons + team_points_l12m | as.factor(race_name):as.factor(race_year),data=.)

stargazer(m1,m2,m3,m4, type='text')
```


What if we break out points by stage points/jersey points?
```{r}
temp = midpoint_df %>%
  group_by(race_name,race_year,team_name, team_points_l12m, grouping) %>%
  summarize(
    stage_ppr_1t = mean(ifelse(stage_number == stage_1t,cumulative_stage_points,NA_integer_),na.rm=T),
    jersey_ppr_1t = mean(ifelse(stage_number == stage_1t,total_jersey_final_points/3,NA_integer_),na.rm=T),
    stage_ppr_3t = mean(ifelse(stage_number == last_stage,cumulative_stage_points,NA_integer_),na.rm=T) - mean(ifelse(stage_number == last_stage - stage_1t & !abandoned,cumulative_stage_points,NA_integer_),na.rm=T),
    jersey_ppr_3t = mean(ifelse(stage_number == last_stage,total_jersey_final_points/3,NA_integer_),na.rm=T),
    riders_1t = sum(ifelse(stage_number == stage_1t,1,0),na.rm=T),
    riders_3t = sum(ifelse(stage_number == last_stage,1,0),na.rm=T)
  ) %>%
  filter(grouping != 'Abandons outside middle third',
         !is.nan(riders_1t)) %>%
  mutate(
    abandons = riders_1t - riders_3t,
    stage_ppr_increase = stage_ppr_3t - stage_ppr_1t,
    jersey_ppr_increase = jersey_ppr_3t - jersey_ppr_1t
  )

m1 = temp %>% felm(stage_ppr_increase ~ abandons,data=.)
m2 = temp %>% felm(stage_ppr_increase ~ abandons + team_points_l12m | as.factor(race_name):as.factor(race_year),data=.)
m3 = temp %>% felm(jersey_ppr_increase ~ abandons,data=.)
m4 = temp %>% felm(jersey_ppr_increase ~ abandons + team_points_l12m | as.factor(race_name):as.factor(race_year),data=.)

stargazer(m1,m2,m3,m4, type='text')
```
Just confused tbh. how does this become a big positive increase per rider??
So close on this! And will be ready to put together

old:
General hypotheses:

* Trains to keep riders at the front do basically nothing, either for GC or sprints
* Mountain domestiques to very little, especially for underdogs
* Attacking in turns is more effective relative to leadouts
* One good leadout man is better than a team of middling leadout men
* Breakaways are relatively underused. EG tim declerq should be going for breaks more often

Talk about pidcock/rodriguez in tour doc. These are oversensationalized but they often contain some truth, like pidcock actually beefing with management. What benefit could he have possibly given carlos??

How can riders use game theory more effectively?

https://www.youtube.com/watch?v=SQ01eBEJ1Tk - 104:30 on unlocking riders

* Fakes:
  * Riders should want opposing riders to not know when they will attack. One way: launch fake attacks, or ride so that lead up to an attack looks just like normal riding
  * Or: some portion of attacks should fizzle early. That way teams will not chase so hard
* 
